---
title: "The Exponential Distribution vs the Central Limit Theorem"
author: "Zoheb Nensey"
---

### Overview
This document is intended to show the differences between the exponential distribution and the distribution that results as a result of the central limit theorem (the normal distribution). 

### Simulations
```{r echo=FALSE, results="hide",message=FALSE}
library(knitr)
library(ggplot2)
library(dplyr)
```

First, I set a number of constants in concurrence with the assignment. For this assignment, the lambda is the rate constant, and is uniform for all simulations. 

```{r results="hide"}
# Setting a seed so that this is reproducible. 
set.seed(234234)

# Setting lamba.
lambda <- 0.2

# Setting number of simulations.
nosim <- 1000

# Number of samples (per assignment)
n <- 40
```

Then, I ran the simulations. 
```{r cache=TRUE}
exponential_distributions <- matrix(rexp(nosim*n, rate=lambda),nosim,n)
exponential_means <- data.frame(means=apply(exponential_distributions, 1, mean))
```
### Comparing the sample mean to the theoretical mean. 
